Order reduction and MILP models for certifying Robustness of DNNs globally.

### 2025-08-02 AAAI2026/Lipitschz.tex
full paper

	\hfill

>>>>

	%\vspace{-0.5cm}
	\input{related}

\section{Conclusion}

In this paper, we introduced a novel "2v" MILP model for global verification questions, which is 4 to 8 times more accurate than the classical MILP model,
and more accurate than the ITNE model of \cite{lipshitz}, see 
Table \ref{table.classical}. 
Using our "3v" or "1v" models abstracting away some features of the "2v" model,
 even better bounds are reached (Tables 3, 4, 5). We also explained how to support $L_1$-perturbations efficiently on top of classical $L_\infty$ perturbations (Prop \ref{prop.l1}). To further lower the bounds by $> 10$ times, we resort to model order reduction, reducing the space of search closer to the space of the training dataset, without sacrificing DNN accuracy ($97\%$). This results into, to the best of our knowledge, the first time real-time robustness certification is reported for a DNN, with $86\%$ of robustness for a $L_1$-perturbation of $.5$ and $53\%$ of robustness for a $L_1$-perturbation twice as large of $1$. Still, global verification is only at its beginning, and combining different methods is necessary to scale to more complex DNNs.

### 2025-08-02 AAAI2026/experiments.tex
full paper

\begin{table}[b!]
	\centering
	\begin{tabular}{|l|c|c|c|}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
	Classical, $50 \times 2$ &    $.320$ &  $.320$ & $.017$ 
    \\\hline
	ITNE, $50 \times 2$ &    $.042$ &  $.037$ & $.022$
	\\ \hline
    2v model, $50 \times 2$ &    {\bf .040} &  $.037$ &  $.018$ 
    \\\hline \hline
    Classical, $100 \times 2$ &  .186  &  $.022$ & $.022$ 
    \\\hline
	ITNE, $100 \times 2$ &    $.045$ &  $.023$ & .023
    \\\hline
	2v model, $100 \times 2$&     {\bf .042} &  $.023$ &   .023 
    \\\hline \hline
	\end{tabular}
	\caption{Comparison of the classical encoding, ITNE and our "2v" model on the pipe system with a fixed timeout of 1000s, with either $50 \times 2$, 
    or the full $100 \times 2$ binary variables.}
    \label{table.classical}
\end{table}

We start by evaluating how accurate the 2v model is vs the classical encoding vs the interleaving twin-network encoding (ITNE) from \cite{lipshitz}, that we reimplement in Gurobi by considering the classical encoding plus the explicit constraints $\frac{x_i-x_i' - \gamma}{2} \leq \hat{x}_i-\hat{x}_i' \leq \frac{x_i-x_i' + \gamma}{2}$, corresponding to Eq. (3) of \cite{lipshitz} (and also to the relaxation of our "1v" model). We compare when half the ReLUs are approximated with the LP relaxation, and when all the ReLUs are encoded exactly. We use the pipe models which is easier to verify, and set the time-out at 1000s for all models.

@@

	\begin{table}[h!]
		\centering
	\begin{tabular}{||l||c|c|c||}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
		1v, $500 \times 1$ & $15.67$ & $.620$ & $.012$ \\\hline 
		3v, $500 \times 3$ & $16.09$ & $.755$ & {\bf .207} \\\hline 
	    2v, $500 \times 2$ & {\bf 8.99} & n/a & n/a \\\hline\hline	 
	\end{tabular}
	\caption{Bounds on $\beta^{.5}_{6,8}$ 
	obtained by the "1v", "3v" and "2v" models 
	on the {\bf full dimension} MNIST DNN, 
	for timeouts of $14400$s, when all 500 ($\times 1$, $\times 2$, $\times 3$) variables are binary.}
	\label{table.mnist}
\end{table}

@@

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.5]{image.png} \hspace{0.8cm}
\includegraphics[scale=0.5]{perturb.png}
\caption{An improbable image and its perturbation (difference at x=13, y=18) 
with maximal $\beta^{.5}_{6,8}=0.207$ for MNIST as obtained by the "3v" $500 \times 2$ model in {\bf full 784 dimension} image space.}
\label{fig3}
\end{figure*}

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.5]{redimage.png} \hspace{0.8cm}
\includegraphics[scale=0.5]{redperturb.png}
\caption{An image and its perturbation from the {\bf 20-dimension reduced space} with maximal $\beta^{.5}_{6,8}=.084$ for MNIST as obtained by the "2v" $500 \times 2$ model.}
\label{fig4}
\end{figure*}

@@

We report in Table \ref{table.reduced} the same $\beta^{.5}_{6,8}$
as in Table \ref{table.mnist}, obtained with the same time-out. The bounds are directly comparable: the best $\beta^{.5}_{6,8}$ obtained using the reduced dimension is $\approx 8$ times smaller than when using the full dimension ($1.137$ vs $8.99$). With these bounds, most images ($86\%$) can be certified in real-time robust for a perturbation $L_1 \leq .5$, and even $53\%$ with a perturbation $L_1 \leq 1$ twice as large, see Table \ref{table.cert}. We illustrate that improbable images are removed by displaying in Fig.~\ref{fig4} the worst-case obtained for $\beta^{.5}_{6,8}$, which indeed looks like a realistic MNIST instance.

@@

>>>>

\begin{table}[b!]
	\centering
	\begin{tabular}{||l|c|c|c||}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
	Classical, $50 \times 2$ &    $.320$ &  $.320$ & $.017$ 
    \\\hline
	ITNE, $50 \times 2$ &    $.042$ &  $.037$ & $.022$
	\\ \hline
    2v model, $50 \times 2$ &    {\bf .040} &  $.037$ &  $.018$ 
    \\\hline \hline
    Classical, $100 \times 2$ &  .186  &  $.022$ & $.022$ 
    \\\hline
	ITNE, $100 \times 2$ &    $.045$ &  $.023$ & .023
    \\\hline
	2v model, $100 \times 2$&     {\bf .042} &  $.023$ &   .023 
    \\\hline \hline
	\end{tabular}
	\caption{Comparison of the classical encoding, ITNE and our "2v" model on the pipe system with a fixed timeout of 1000s, with either $50 \times 2$, 
    or the full $100 \times 2$ binary variables.}
    \label{table.classical}
\end{table}

We start by evaluating how accurate the 2v model is vs the classical encoding vs the interleaving twin-network encoding (ITNE) from \cite{lipshitz}, that we reimplement in Gurobi by considering the classical encoding plus the explicit constraints $\frac{x_i-x_i' - \gamma}{2} \leq \hat{x}_i-\hat{x}_i' \leq \frac{x_i-x_i' + \gamma}{2}$, corresponding to Eq. (3) of \cite{lipshitz} (and also to our (\ref{eq.lpr})). We compare when half the ReLUs are approximated with the LP relaxation, and when all the ReLUs are encoded exactly. We use the pipe models which is easier to verify, and set the time-out at 1000s for all models.

@@

	\begin{table}[h!]
		\centering
	\begin{tabular}{||l||c|c|c||}\hline\hline
		model &        Bound $\downarrow$ &  Sol. &      Worst-Case $\uparrow$ \\\hline \hline
		1v, $500 \times 1$ & {\bf 14.97} & $.845$ & $.009$ \\\hline 
		3v, $500 \times 3$ & $17.66$ & $.813$ & {\bf .518} \\\hline 
	    2v, $500 \times 2$ & $16.49$ & n/a & n/a \\\hline\hline	 
	\end{tabular}
	\caption{Bounds on $\beta^{.5}_{6,8}$ 
	obtained by the "1v", "3v" and "2v" models 
	on the {\bf full dimension} MNIST DNN, 
	for timeouts of $14400$s, when all 500 ($\times 1$, $\times 2$, $\times 3$) variables are binary.}
	\label{table.mnist}
\end{table}

@@

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.6]{image.png} \hspace{1.5cm}
\includegraphics[scale=0.6]{perturb.png}
\caption{An improbable image and its perturbation ($L1$-difference of $.5$ in a unique pixel at x=14, y=8 from top) with maximal $\beta^{.5}_{6,8}=0.518$ for MNIST as obtained by the "3v" $500 \times 2$ model in {\bf full 784 dimension} image space.}
\label{fig3}
\end{figure*}

\begin{figure*}[t!]
	\centering
\includegraphics[scale=0.5]{redimage.png} \hspace{1.5cm}
\includegraphics[scale=0.5]{redperturb.png}
\caption{An image and its perturbation from the {\bf 20-dimension reduced space} with maximal $\beta^{.5}_{6,8}=.084$ for MNIST as obtained by the "2v" $500 \times 2$ model.}
\label{fig4}
\end{figure*}

@@

We report in Table \ref{table.reduced} the same $\beta^{.5}_{6,8}$
as in Table \ref{table.mnist}, obtained with the same time-out. The bounds are directly comparable: the best $\beta^{.5}_{6,8}$ obtained using the reduced dimension is $>10$ times smaller than when using the full dimension ($1.137$ vs $14.97$). 
With these bounds, most images ($86\%$) can be certified in real-time robust for a perturbation $L_1 \leq .5$, and even $53\%$ with a perturbation $L_1 \leq 1$ twice as large, see Table \ref{table.cert}. We illustrate that improbable images are removed by displaying in Fig.~\ref{fig4} the worst-case obtained for $\beta^{.5}_{6,8}$, which indeed looks like a realistic MNIST instance.

@@

\newpage

\noindent {\bf Supplementary material content:} We provide in supplementary materials additional content, in particular results with reduced number of binary variables. We also provide the proof of Prop.~\ref{Prop2}, and explanations on the reduced-order dimension pipeline.

### 2025-08-02 AAAI2026/related.tex
full paper

Some works CITE
Leino et al. [2021]; Zhang et al. [2022a] Sun et al. [2022]  Chen et al. [2021], REGLO focus on training a network to be - rather than proving that the DNN is as we do - gobally robust .
Other work estimate global robustness bounds, but without hard verification certificates CITE Bastani et al. [2016]; Ruan et al. [2019], Gopinath et al. [2018], e.g. using probabilistic guarantees  Levy et al. [2023]; Mangal et al. [2019].

Closer to our paper, several works consider a kind of global Lipschitz bound, to understand how the output change with input perturbations, as \cite{Marabou} and \cite{ITNE,GROCET}. However, they do not consider whether the classification changes.

@@

Technically, none of the MILP base work \cite{vhagar,ITNE} 
consider $L_1$ bounds. Further, they use the classical MILP encoding from \cite{MILP}, and not our novel "2v" model, nor even our "1v" model. However, \cite{ITNE} 
defines the {\em diff variables} $y,\hat{y}$, adding explicitely the linear relaxation $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$ implied by our models, Eq. (3) in \cite{ITNE} corresponding essentially to it. \cite{vhagar} does things differently, adding per neuron constraints (depending on the perturbation) between the (binary) variables to simplify the MILP problem.

>>>>

Some works \cite{Leino,Zhang22a,Sun22,Chen21,REGLO} focus on training a network to be - rather than proving the DNN is - globally robust.
Other works \cite{Bastani16,Ruan19,Gopinath18} estimate global robustness bounds, but without hard verification certificates, e.g. using probabilistic guarantees 
\cite{Levy23,Mangal19}.

Closer to our paper, several works  \cite{Marabou,lipshitz,GROCET} consider a global certified restricted Lipschitz bound, to understand how the output change with input perturbations. However, they do not consider whether the classification changes.

@@

Technically, none of the MILP base work \cite{vhagar,ITNE} 
consider $L_1$ bounds. Further, they use the classical MILP encoding from \cite{MILP}, and not our novel "2v" model, nor even our "1v" model. However, \cite{ITNE} 
defines the {\em diff variables} $y,\hat{y}$, adding explicitely the linear relaxation $\frac{y_i-\gamma_i}{2} \leq \hat{y}_i \leq \frac{y_i+\gamma_i}{2}$ implied by our models, Eq. (3) in \cite{ITNE} corresponding essentially to it. \cite{vhagar} does things differently, adding per neuron constraints (depending on the perturbation) between the (binary) variables to simplify the MILP problem, which is well-adapted to occlusion and patches but less to $L_1$ perturbations.
